---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(digits = 0)

require(rnflstats)
require(magrittr)
require(grid)
```

This R package will eventually simplify a multitude of tasks when analyzing NFL play-by-play data. In its current iteration, it is based on the [New York Times 4th Down Bot](http://nyt4thdownbot.com/) ([Github repository](https://github.com/TheUpshot/4thdownbot-model)) and performs tasks limited to evaluating decisionmaking on 4th down plays. Using NFL play-by-play data from [Armchair Analysis](http://armchairanalysis.com/), this code will:

- Munge the raw play-by-play data and transform it into a form suitable for modeling
- Create a win probability model and serialize that model
- Provide all of the functions to make optimal 4th down decisions for a given play

The Armchair Analysis data is *not free*. It costs $49 (one-time) to gain access to play-by-play data from the 2000-2014 NFL seasons. There is also a professional package that will provide weekly updates. Without the Armchair Analysis data, you will not be able to use much of this code.

# Usage
Unzip the play-by-play data into a directory. Run the following code from the directory where you want the Fourth Down Bot code to live. It will create the subdirectories `models` and `data` to store files. The significant difference from the NYT 4th Down Bot is a more basic field goal success prediction model (still a work-in-progress). 

```{r results = 'hide', error = FALSE, message = FALSE, warning = FALSE}
data_prep("data/armchair_analysis/")
model_train()
fg_model_train()
```

If you wish to view the calibration plots and ROC curves for the model, run `model_train` or `fg_model_train` with the `plot = TRUE` flag, like so (as you will see, the field goal model needs a lot more work to even match the current NYT model):

```{r error = FALSE, message = FALSE, warning = FALSE}
model_train(plot = TRUE)
fg_model_train(plot = TRUE)
```

To evaluate the optimal decision, you must first load the already compiled data into your R workspace, then run the `run_bot` function. Currently you can only input your key information interactively via text prompts in R. Alternatively, for testing purposes you can use the `random_play` function to randomly generate fake data to test the package's functions. I am currently working on a better way to return summary information on this decision.

```{r error = FALSE, message = FALSE, warning = FALSE}
nfl_data <- load_data()
results <- run_bot(nfl_data$data, nfl_data$win_model, nfl_data$fg_model, random_play(nfl_data$data))
plot_decision(results) %>%
  grid.draw
```

| Option | Chance of Converting | Before Play | After Play | Change |
|:-------|:---------------------|:------------|:-----------|:-------|
| Go for it | `r results$decision$prob_success * 100`% | `r results$probs$pre_play_wp * 100`% | `r results$probs$wp_ev_goforit * 100`% | `r (results$probs$wp_ev_goforit - results$probs$pre_play_wp) * 100`% |
| Punt | -- | `r results$probs$pre_play_wp * 100`% | `r results$probs$punt_wp * 100`% | `r (results$probs$punt_wp - results$probs$pre_play_wp) * 100`% |
| Field goal try | `r results$probs$prob_success_fg * 100`% | `r results$probs$pre_play_wp * 100`% | `r results$probs$fg_ev_wp * 100`% | `r (results$probs$fg_ev_wp - results$probs$pre_play_wp) * 100`% |



